---
title: "3. Polygons. Simplify The UK."
output:
  rmarkdown::html_document:
    code_folding: hide
    theme: sandstone   
    highlight: tango
    css: "css/andrewl.css"
    includes:
      before_body: "partials/header.html"
      after_body: "partials/footer.html"
---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> Want to see the code? Click on the black boxes on the right to show/hide the code. 

Simply speaking a polygon is a collection of lines that join together to produce
a closed shape. 

All polygons are simplified versions of the real world. But we can simplify them
further.

In this map we take the UK and simplify it to extreme levels. We're going to 
increase the maximum distance between points around the UK to 10km, 20km, 50km
and then 100km, although we'll ensure that we don't remove any smaller polygons
(islands)
```{r data, message=FALSE, warning=FALSE, results='show'}
#First, we'll need to load a bunch of libraries so we can handle and view geospatial data
library('sf')
library('mapview')

# We can use the st_read function from the sf library to read in the data from the geopackage
# and transform it to the British National Grid (EPSG:27700)
# This is done so that the data is in metres and not degrees
countries <- st_read('./data/03/geoBoundaries-GBR-ADM1_simplified.shp', quiet = TRUE) %>%
st_transform(countries, crs = 27700)

# Create 4 data frames with simplified topologies at various tolerances
s10000 <- st_simplify(countries,  preserveTopology = TRUE, dTolerance = 10000)
s20000 <- st_simplify(countries,  preserveTopology = TRUE, dTolerance = 20000)
s50000 <- st_simplify(countries,  preserveTopology = TRUE, dTolerance = 50000)
s100000 <- st_simplify(countries,  preserveTopology = TRUE, dTolerance = 100000)


mapviewOptions(basemaps = c("Esri.WorldShadedRelief"),
               layers.control.pos = "topright",
               leafletWidth = "100%",
               legend = FALSE
               )

# Add the layers to a map
map  = mapview(s100000,  layer.name = "100km", col.regions = "yellow", zcol="shapeName") +
    mapview(s50000,  layer.name = "50km", col.regions = "blue", zcol="shapeName") +
    mapview(s20000, layer.name = "20km", col.regions = "green", zcol="shapeName") +
  mapview(s10000, layer.name= "10km", col.regions = "red", zcol="shapeName") 

```

### The Map

Here we have the UK simplified to 10,000, 20,000, 50,000 and 100,000 metres. You
can use the layer control in the top-left to turn on/off individual layers.
```{r map, results='show'}
#Display the map
map
```

### Credits

Data: For this map we use data from [geoBoundaries](https://www.geoboundaries.org), specifically
the Simplified Country Files for the UK found at https://www.geoboundaries.org/simplifiedDownloads.html
