---
title: "4. Hexagons: How Uber sees the world"
output:
  rmarkdown::html_document:
    code_folding: hide
    theme: sandstone   
    highlight: tango
    css: "css/andrewl.css"
    includes:
      before_body: "partials/header.html"
      after_body: "partials/footer.html"

---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> Want to see the code? Click on the black boxes on the right to show/hide the code. 

Uber have devised a grid system for the world, using hexagons, called h3. Comprised of a number of layers, the hexagons at each
layer are of different sizes (so the more you zoom in, the smaller the hexagons get). Given a location and a layer, you can
find the hexagon that contains that location.

Why hexagons?

* They tessellate (fit together without gaps)
* Unlike squares or triangles they tessalate on edges rather than corners
* This means you can add properties to the edges of the hexagons to indicate things like how 'expensive' it to move between two hexagons
* This also means the distance between the centre of a hexagon and the centre of its neighbours is the same in all directions
* You can approximate a circle with a hexagon or a ring of hexagons


H3 is open source and there are implementations in a number of languages. You can find out more about it [here](https://eng.uber.com/h3/).


### The Map

This is a map of the UK overlaid with hexagons at different layers. The hexagons are coloured by layer. As you zoom in, the hexagons get
smaller. Hover over a hexagon to see its h3 index.

```{r map, message=FALSE, warning=FALSE, results='show'}
library('h3')
library('sf')
library('leaflet')

# Let's reuse the Ordnance Survey populated places data from an earlier map
placesOSGB <- st_read('./data/02/opname_gb.gpkg', quiet = TRUE, query='SELECT geometry, name1 as placeName from named_place where type = "populatedPlace"')

# Convert to WGS84
places <- st_transform(placesOSGB, 4326 )

# We'll create hexagons at different layers based on the places
h3_index_02 <- geo_to_h3(places, 2)
tbl <- table(h3_index_02) %>%
  tibble::as_tibble()
h302 <- h3_to_geo_boundary_sf(tbl$h3_index_02) 

h3_index_03 <- geo_to_h3(places, 3)
tbl <- table(h3_index_03) %>%
  tibble::as_tibble()
h303 <- h3_to_geo_boundary_sf(tbl$h3_index_03) 

h3_index_04 <- geo_to_h3(places, 4)
tbl <- table(h3_index_04) %>%
  tibble::as_tibble()
h304 <- h3_to_geo_boundary_sf(tbl$h3_index_04)

h3_index_05 <- geo_to_h3(places, 5)
tbl <- table(h3_index_05) %>%
  tibble::as_tibble()
h305 <- h3_to_geo_boundary_sf(tbl$h3_index_05)

h3_index_06 <- geo_to_h3(places, 6)
tbl <- table(h3_index_06) %>%
  tibble::as_tibble()
h306 <- h3_to_geo_boundary_sf(tbl$h3_index_06)

h3_index_07 <- geo_to_h3(places, 7)
tbl <- table(h3_index_07) %>%
  tibble::as_tibble()
h307 <- h3_to_geo_boundary_sf(tbl$h3_index_07)

h3_index_08 <- geo_to_h3(places, 8)
tbl <- table(h3_index_08) %>%
  tibble::as_tibble()
h308 <- h3_to_geo_boundary_sf(tbl$h3_index_08)

# Render the hexagons on a leaflet map
map <- leaflet(width = "100%") %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addPolygons(
    data = h303,
    weight = 2,
    color = "red",
    fillColor = "red",
    fillOpacity = 0.4,
    label = ~ sprintf("%s", h3_index),
    group = "H303"
  ) %>%
  groupOptions("H303", zoomLevels = 04:05) %>%
  addPolygons(
    data = h304,
    weight = 2,
    color = "blue",
    fillColor = "blue",
    fillOpacity = 0.4,
    label = ~ sprintf("%s", h3_index),
    group = "H304"
  ) %>%
  groupOptions("H304", zoomLevels = 06:07) %>%
  addPolygons(
    data = h305,
    weight = 2,
    color = "green",
    fillColor = "green",
    fillOpacity = 0.4,
    label = ~ sprintf("%s", h3_index),
    group = "H305"
  ) %>%
  groupOptions("H305", zoomLevels = 08:09) %>%
  addPolygons(
    data = h306,
    weight = 2,
    color = "purple",
    fillColor = "purple",
    fillOpacity = 0.4,
    label = ~ sprintf("%s", h3_index),
    group = "H306"
  ) %>%
  groupOptions("H306", zoomLevels = 10:11) %>%
  addPolygons(
    data = h307,
    weight = 2,
    color = "orange",
    fillColor = "orange",
    fillOpacity = 0.4,
    label = ~ sprintf("%s", h3_index),
    group = "H307"
  ) %>%
  groupOptions("H307", zoomLevels = 12:13) %>%
  addPolygons(
    data = h308,
    weight = 2,
    color = "black",
    fillColor = "black",
    fillOpacity = 0.4,
    label = ~ sprintf("%s", h3_index),
    group = "H308"
  ) %>%
  groupOptions("H308", zoomLevels = 14:15)


map

```


### Credits

H3: https://eng.uber.com/h3/

Data: Ordnance Survey Open Places OGL https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/
