---
title: "5. A Journey - Up Snowdon/Yr Wyddfa"
output:
  rmarkdown::html_document:
    code_folding: hide
    theme: sandstone   
    highlight: tango
    css: "css/andrewl.css"
    includes:
      before_body: "partials/header.html"
      after_body: "partials/footer.html"

---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> Want to see the code? Click on the black boxes on the right to show/hide the code. 

### Snowdon / Yr Wyddfa

The highest peak in Wales, higher than any in England. There are many ways to
walk up it - I recommend walking up from the quieter Beddgelert side - Snowdon
Ranger or Rhyd Ddu paths are usually attract far fewer visitors and are much
more peaceful than the others.  
 
### The Map

A really simple map today. More than one journey though - these are some of the
routes I've used to walk up Snowdon/Yr Wyddfa. Taken from gpx files that I've
downloaded from my Strava account and drawn over OpenStreetMap data in the
OpenTopoMap style.

```{r map, message=FALSE, warning=FALSE, results='show'}
#First, we'll need to load a bunch of libraries so we can handle and view geospatial data
library('sf')
library('mapview')

# We can use the st_read function from the sf library to read in the data from 
# the GPX files we have. We must specify the 'tracks' layer as this 
s1 <- st_read('./data/05/01.gpx', layer = "tracks", quiet = TRUE)
s2 <- st_read('./data/05/02.gpx', layer = "tracks", quiet = TRUE)
s3 <- st_read('./data/05/03.gpx', layer = "tracks", quiet = TRUE)
s4 <- st_read('./data/05/04.gpx', layer = "tracks", quiet = TRUE)
s5 <- st_read('./data/05/05.gpx', layer = "tracks", quiet = TRUE)


mapviewOptions(basemaps = c("OpenTopoMap"),
               leafletWidth = "100%",
               legend = FALSE,
               homebutton = FALSE,
               )

# Finally plot the layers on the map in different colours.
mapview(s1, color = "red", lwd=6, layer.name="Watkin/Miners") + 
mapview(s2, color = "green", lwd=6, layer.name="Snowdon Ranger") +
mapview(s3, color = "purple", lwd=6, layer.name="South Ridge/Pyg") + 
mapview(s4, color = "blue", lwd=6, layer.name="Rhyd Ddu/Snowdon Ranger/Quarries") + 
mapview(s5, color = "brown", lwd=6, layer.name="Miners/Snowdon Ranger")
```
