---
title: "13. A New Tool - Is It In London?"
output:
  rmarkdown::html_document:
    code_folding: hide
    theme: sandstone   
    highlight: tango
    css: "css/andrewl.css"
    includes:
      before_body: "partials/header.html"
      after_body: "partials/footer.html"
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> Want to see the code? Click on the black boxes on the right to show/hide the code. 
 
Is it in London?

There are some places that lie within the M25 but are not in any London Borough. TFL might operate some buses in these areas,
you might be able to use a London Underground station, but you won't be able to vote in London elections. Depending on your
point of view (usually where you live), these places might be 'in London' or they might not.

This map shows the area within the M25 that is not part of any London Borough. To build this, I used a bunch of new geometry tools
and:

- Used the M25 and A282 roads, noting that the M25 doesn’t fully circle London (the Dartford Crossing is designated as the A282).

- Joined the M25 and A282 roads to form a continuous path.

- Created a convex hull around this path to generate an M25 polygon.

- Loaded a London Wards shapefile to build a polygon representing London.

- Subtracted the London polygon from the M25 polygon to isolate the area within the M25 but outside London.
   
- Created a rectangle covering the UK and subtracted the M25 polygon from it, further refining the area to only within the M25 and outside London.


```{r map, message=FALSE, warning=FALSE, results='show'}
# Let's loadng the sf and mapview libraries to process and view geospatial data
library(sf)
library(leaflet)
library(raster)
library(terra)
library(concaveman)

# Load the M25 and A282 roads
m25 <- st_read('./data/13/Data/oproad_gb.gpkg', quiet = TRUE, query = 'select * from road_link where (road_classification_number = "M25" or road_classification_number = "A282")')

# Load the London Wards shapefile
london_wards <- st_read('./data/13/London-wards-2018_ESRI/London_Ward.shp', quiet = TRUE)

# Combine the M25 and A282 roads to form a continuous path
m25_coords <- st_union(m25) %>% st_cast('POINT') %>% st_union() %>% st_coordinates()

# Use concaveman to create a concave hull around the points
polygon_boundary <- concaveman(m25_coords)

# Convert the concave hull back to an sf polygon
m25_poly <- st_sfc(st_polygon(list(polygon_boundary)), crs = 27700)

# Subtract the London Wards from the M25 polygon
inside_m25_outside_london <- st_difference(m25_poly, st_union(london_wards))

# The bounds of the UK in EPSG:27700
minx <- 0
miny <- 0
maxx <- 700000
maxy <- 1300000

# Create a rectangle vector
rect <- st_sfc(st_polygon(list(rbind(c(minx, miny), c(maxx, miny), c(maxx, maxy), c(minx, maxy), c(minx, miny)))), crs=27700)

# Punch a hole in the rectangle using the M25 polygon
m25_hole <- st_difference(rect, inside_m25_outside_london) %>% st_transform(crs=4326)

# Create a leaflet map
map <- leaflet(width = "100%") %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addPolygons(
    data = m25_hole,
    color = "white",
    fillColor = "white",
    fillOpacity = 0.95
  )%>%
  setView(lng = -0.1407376, lat = 51.4908968, zoom = 10) 

map

```
### Credits

Data: Ordnance Survey Open Roads from https://osdatahub.os.uk/downloads/open/OpenRoads OGLv3 https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/

London Wards Boundaries from https://data.london.gov.uk/dataset/statistical-gis-boundary-files-london OGLv2 https://www.nationalarchives.gov.uk/doc/open-government-licence/version/2/
