---
title: "15. My Data - Where Are They Flying To?"
output:
  rmarkdown::html_document:
    code_folding: hide
    theme: sandstone   
    highlight: tango
    css: "css/andrewl.css"
    includes:
      before_body: "partials/header.html"
      after_body: "partials/footer.html"
---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> Want to see the code? Click on the black boxes on the right to show/hide the code. 

A heatmap of the destinations of the aircraft that fly over my house. I've rigged up
an [ADSB](https://en.wikipedia.org/wiki/Automatic_Dependent_Surveillance%E2%80%93Broadcast) receiver to
a Raspberry Pi and some software. 

The software listens for any aircraft that are overhead - the aircraft broadcast their identifiers, altitude, speed etc.

Over time it's collected a lot of data about where these aircraft were heading to, and present them on this map.

### The Map

The map shows the destinations of aircraft that have flown over my house - the size of the circles
are proportional to the number of aircraft that were destined to land at that location and the colour
represents the average altitude of the aircraft when they overflew.

```{r map, message=FALSE, warning=FALSE, results='show'}
#First, we'll need to load a bunch of libraries so we can handle and view geospatial data
library(sf)
library(mapview)
library(jsonlite)

jsonData <- data.frame(fromJSON('./data/15/watft_destinations.json'))

colnames(jsonData) <- c("lat", "lng", "count", "Average Altitude")

destinations <- st_as_sf(jsonData, coords = c("lat", "lng"), 
                                   crs = 4326)

mapviewOptions(basemaps = c("Esri.WorldTopoMap"),
               layers.control.pos = "topright",
               legend.pos = "bottomleft",
               leafletWidth = "100%",
               legend = TRUE
               )

pal = mapViewPalette("Pastel 1")

mapview(destinations, zcol="Average Altitude", cex="count",  col.regions=RColorBrewer::brewer.pal(7, "Reds"))
``` 


### Credits

Data: Ordnance Survey Open Places OGL https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/
