---
title: 16. Choropleth - World Population Density
output:
  rmarkdown::html_document:
    code_folding: hide
    theme: sandstone   
    highlight: tango
    css: "css/andrewl.css"
    includes:
      before_body: "partials/header.html"
      after_body: "partials/footer.html"
---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> Want to see the code? Click on the black boxes on the right to show/hide the code. 

## The Map

A choropleth map is a type of thematic map in which areas are shaded or patterned in proportion to the value of a statistical variable being represented on the map.

This simple map shows the population density of each country in the world. The darker the color, the higher the population density. 

```{r map, message=FALSE, warning=FALSE, results='show'}
#First, we'll need to load a bunch of libraries so we can handle and view geospatial data
library(sf)
library(mapview)

# Now we'll load the world borders data 
wrld_simpl <- st_read('./data/11/TM_WORLD_BORDERS-0.2.shp', quiet = TRUE) %>% st_transform(crs = 3857)

# Let's calculate the real area of each country
wrld_simpl$realarea =  as.numeric(st_area(wrld_simpl))

# and calculate the population density of each country
wrld_simpl$PopulationDensity <- (wrld_simpl$POP2005 / wrld_simpl$realarea) * 10000

mapviewOptions(basemaps = c("CartoDB.Positron"),
               layers.control.pos = "topright",
               legend.pos = "bottomleft",
               leafletWidth = "100%",
               legend = FALSE)

# Now we can plot the map, note that we're using custom breaks and colors
mapview(wrld_simpl, zcol="PopulationDensity", label="NAME", at=c(0,0.2,0.5,1,2,5,10,20), col.regions=RColorBrewer::brewer.pal(5, "Reds"))

```

## Credits

Map data from the World Wind Java project: https://github.com/nasa/World-Wind-Java/blob/master/WorldWind/testData/shapefiles/TM_WORLD_BORDERS-0.2Readme.txt
