---
title: "18. 3D - Routes of Snowdon"
output:
  rmarkdown::html_document:
    code_folding: hide
    theme: sandstone   
    highlight: tango
    css: "css/andrewl.css"
    includes:
      before_body: "partials/header.html"
      after_body: "partials/footer.html"

---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> Want to see the code? Click on the black boxes on the right to show/hide the code. 

### Data

An extension of [05 A Journey](05-a-journey.html) but this time the map is in 3D. I've taken
3D elevation data from the SRTM dataset, overlaid with OpenStreetMap data for roads and waterways.
The routes are from my GPX files downloaded from Strava. I've added shadows and textures to the
3D map to make it look more realistic.

I used Tyler Morgan-Wall's `rayshader` package to create the 3D map, and used his blog posts at
[tylermw.com](https://www.tylermw.com) for inspiration.

I've used the `rayshader` package to render the 3D map at various angles and then created a video
to show it off in all its glory.


### The Map

```{r data, message=FALSE, warning=FALSE, results='show'}
library(rayshader)
library(sp)
library(raster)
library(scales)
library(sf)
library(osmdata)
library(ggplot2)
library(dplyr)

# Load the elevation data
snowdon_elevation <- raster::raster("./data/18/output_SRTMGL1.tif")
snowdon_elevation_matrix <- rayshader::raster_to_matrix(snowdon_elevation)

# Set the extent of the elevation data
osm_bbox = c(-4.23541666664337,53.0037499999991,-3.96347222219889,53.127638888888)

# Get the OpenStreetMap data for roads and waterways
roads = opq(osm_bbox) %>% 
  add_osm_feature("highway") %>% 
  osmdata_sf() 

water = opq(osm_bbox) %>% 
  add_osm_feature("water") %>% 
  osmdata_sf() 

waterway = opq(osm_bbox) %>% 
  add_osm_feature("waterway") %>% 
  osmdata_sf() 

# Get coordinates of the routes
s1 <- st_read("./data/05/01.gpx", layer = "tracks", quiet = TRUE) %>% st_transform(crs = crs(snowdon_elevation))
s2 <- st_read("./data/05/02.gpx", layer = "tracks", quiet = TRUE) %>% st_transform(crs = crs(snowdon_elevation))
s3 <- st_read("./data/05/03.gpx", layer = "tracks", quiet = TRUE) %>% st_transform(crs = crs(snowdon_elevation))
s4 <- st_read("./data/05/04.gpx", layer = "tracks", quiet = TRUE) %>% st_transform(crs = crs(snowdon_elevation))
s5 <- st_read("./data/05/05.gpx", layer = "tracks", quiet = TRUE) %>% st_transform(crs = crs(snowdon_elevation))

# Combine the routes into a single dataframe
routes <- dplyr::bind_rows(list(s1,s2,s3,s4,s5))

# Transform the roads and waterways to the same CRS as the elevation data
roads_lines = st_transform(roads$osm_lines, crs=crs(snowdon_elevation)) %>% filter(highway %in% c("primary", "secondary", "tertiary"))
water_lines = st_transform(waterway$osm_lines, crs=crs(snowdon_elevation))
water_polys = st_transform(water$osm_polygons, crs=crs(snowdon_elevation))

# Generate an overlay for the routes, roads and waterways
route_line_overlay = generate_line_overlay(routes, heightmap = snowdon_elevation_matrix, extent=extent(snowdon_elevation), linewidth = 6, color="yellow")
route_line_overlay_inner = generate_line_overlay(routes, heightmap = snowdon_elevation_matrix, extent=extent(snowdon_elevation), linewidth = 2, color="red") 

roads_line_overlay = generate_line_overlay(roads_lines, heightmap = snowdon_elevation_matrix, extent=extent(snowdon_elevation), linewidth = 6, color="black") 
roads_line_overlay_inner = generate_line_overlay(roads_lines, heightmap = snowdon_elevation_matrix, extent=extent(snowdon_elevation), linewidth = 4, color="lightgrey") 

water_line_overlay = generate_line_overlay(water_lines, heightmap = snowdon_elevation_matrix, extent=extent(snowdon_elevation), linewidth = 1, color="skyblue2") 

water_poly_overlay = generate_polygon_overlay(water_polys, heightmap = snowdon_elevation_matrix, extent=extent(snowdon_elevation), linewidth = 1, linecolor="skyblue2", palette="skyblue2") 

# Plot the 3D map
base_map = snowdon_elevation_matrix %>% 
  height_shade() %>% 
  add_overlay(sphere_shade(snowdon_elevation_matrix, texture = "desert", zscale=4, colorintensity = 5), alphalayer=0.5) %>%
  add_overlay(water_poly_overlay) %>%
  add_overlay(water_line_overlay) %>%
  add_overlay(roads_line_overlay) %>%
  add_overlay(roads_line_overlay_inner) %>%
  add_overlay(route_line_overlay) %>%
  add_overlay(route_line_overlay_inner) %>%
  add_shadow(lamb_shade(snowdon_elevation_matrix,zscale = 40),0) %>%
  add_shadow(texture_shade(snowdon_elevation_matrix,detail=8/10,contrast=8,brightness = 11), 0.1) %>%
  plot_3d(heightmap=snowdon_elevation_matrix, windowsize = c(1200,800), zscale=40)

# Save the 3D map at various angles
#angles= seq(0,360,length.out = 1441)[-1]
#for(i in 1:1440) {
#  render_camera(theta=-45+angles[i], phi=20, zoom=0.2,  fov=60)

#  render_snapshot(filename = sprintf("snowdon%i.png", i), 
#                  title_text = "Snowdon routes",
#                  title_bar_color = "#1f5214", title_color = "white", title_bar_alpha = 1)
#}
#rgl::rgl.close()

# Create a video of the 3D map from the frames
#system("ffmpeg -framerate 60 -i snowdon%d.png -pix_fmt yuv420p images/snowdon.mp4")
```

![](images/snowdon.mp4)


### Credits

Data: [SRTM](https://www2.jpl.nasa.gov/srtm/),
[OpenStreetMap](https://www.openstreetmap.org/)
