---
title: "22. 2 Colours - Black and White"
output:
  rmarkdown::html_document:
    code_folding: hide
    theme: sandstone   
    highlight: tango
    css: "css/andrewl.css"
    includes:
      before_body: "partials/header.html"
      after_body: "partials/footer.html"
---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> Want to see the code? Click on the black boxes on the right to show/hide the code. 
 

### Map

Here's a map of the UK split into two colours - black and white. The map shows the Voronoi polygons of all the places in the UK with "black" or "white" in their name. Anywhere within a black area is nearer to a town with "black" in it's name than "white", vice-versa
with white areas.

```{r data, message=FALSE, warning=FALSE, results='show'}
#First, we'll need to load a bunch of libraries so we can handle and view geospatial data
library('sf')
library('raster')
library('mapview')
library('terra')
library(dplyr)

#Next, we'll load the data we need from the OS Open Names dataset
blackAndWhite <- st_read('./data/02/opname_gb.gpkg', quiet = TRUE, query='SELECT geometry, name1 as placeName from named_place where type = "populatedPlace" and (name1 like "%black%" or name1 like "%white%")') 

#We'll also need the UK boundary to crop the Voronoi polygons to the UK
uk_boundary <- st_read('./data/11/TM_WORLD_BORDERS-0.2.shp', quiet = TRUE) %>% st_transform(crs = 27700) %>% filter(NAME == "United Kingdom")

#Now we can create the Voronoi polygons
v <- blackAndWhite %>% 
  st_union() %>%
  st_voronoi(envelope=st_as_sfc(uk_boundary)) 

#And crop them to the UK
cropped_v <- st_intersection(st_cast(v), st_union(uk_boundary)) 

#Join the Voronoi polygons to the original data so we can colour them
vv = st_join(st_as_sf(cropped_v), blackAndWhite)

#Add a column to the data to colour the polygons
vv$colour <- ifelse(grepl("lack", vv$placeName), 0, 1)

#Now we can plot the map
mapviewOptions(basemaps = c("Esri.WorldGrayCanvas"),
               leafletWidth = "100%",
               legend = FALSE,
               homebutton = FALSE,
               fgb = FALSE
               )



mapview(vv, label="placeName", zcol="colour", lwd=1, col.regions=list("black","white"),col=list("black"), alpha.regions = 1, aplha = 1 )
```

### Credits

Data: Ordnance Survey Open Places OGL https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/
World Boundaries: WorldWind https://worldwind.arc.nasa.gov/
