---
title: "27. Micromapping"
output:
  rmarkdown::html_document:
    code_folding: hide
    theme: sandstone   
    highlight: tango
    css: "css/andrewl.css"
    includes:
      before_body: "partials/header.html"
      after_body: "partials/footer.html"

---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> Want to see the code? Click on the black boxes on the right to show/hide the code. 

### A superficial map

Here's an animation based on a dataset from the British Geological Survey. The height
of each hexagon represents how deep below the surface the natural bedrock is. Geologists
call this non-rock material 'superficial', and it's been laid down over the past
2.5 million years or so.

I've taken the area around the Bristol channel and animated the superficial
layer falling away. It's a terrible map as it has no scale, north point, the animation
is entirely gratuitous, and it's been rendered on an underpowered machine! But as usual
the code can act as a starting point for building your own maps or exploring this
data.

Why's this under the micromapping category? Well despite covering a large area the 
maximum depth of the Superficial layer on this map is just 20 metres or so - micro in
comparison to the area.

Like my previous 3D attempt, much credit goes to https://www.tylermw.com for his
ggplot library and inspirational blog posts

### The Map 

```{r data, message=FALSE, warning=FALSE, results='show'}
library(rayshader)
library(sp)
library(sf)
library(ggplot2)
library(paletteer)

hex <- read_sf("./data/27/SDTM_HEX/Data/ESRI/BGS_SDTM_1km.shp")

distance <- 50000

box = c(xmin = 328337-distance, ymin = 172855-distance, xmax = 328337+distance, ymax = 172855+distance)
sf::sf_use_s2(FALSE)
a <- st_crop(hex, box)
sf::sf_use_s2(TRUE)

max_depth = max(a$BSTM_MEAN)

p = paletteer_c("ggthemes::Blue-Green Sequential", n=max(a$BSTM_MEAN))

for(i in 0:max(a$BSTM_MEAN)) {

  a$depth = pmax(max_depth-i,max_depth-a$BSTM_MEAN)
  #these ensure we have a consistent height scale across the frames
  a$depth[0] = max_depth
  a$depth[1] = 0

m = ggplot(a) +
  geom_sf(color = "white", aes(colour="white", fill = a$depth )) +
  scale_fill_paletteer_binned("ggthemes::Brown", direction=-1, breaks=0:max(a$BSTM_MEAN)+1) +
  theme_void()  +
  theme(legend.position = "none", axis.line = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_blank(), panel.background = element_rect(fill = "white"), )

plot_gg(m, scale = 50, height_aes = "fill", multicore = TRUE, windowsize = c(1200, 960),
        fov = 75, zoom = 0.3, theta = 0+(i*2), phi = 30, invert=FALSE, shadow = TRUE, raytrace = TRUE, shadow_intensity=0.5, background="grey")

render_highquality(filename=sprintf("/tmp/depth-%d", i), clear = TRUE)

}

for(i in 1:60) {

  plot_gg(m, scale = 50, height_aes = "fill", multicore = TRUE, windowsize = c(1200, 960),
        fov = 75, zoom = 0.3, theta = 0+(max(a$BSTM_MEAN)*2), phi = 30+i, invert=FALSE, shadow = TRUE, raytrace = TRUE, shadow_intensity=0.5, background="grey")

  render_highquality(filename=sprintf("/tmp/depth-%d", i), clear = TRUE)
}


# Create a video of the 3D map from the frames
system("ffmpeg -y -framerate 15 -i /tmp/depth-%d.png -pix_fmt yuv420p images/micromapping.mp4")


```


```

![](images/micromapping.mp4)


### Credits

Data: Superficial thickness 1 km hex grid by the British Geological Survey under OGL license from
https://www.bgs.ac.uk/datasets/superficial-thickness-model-1-km-hex-grid/