---
title: "28. The Blue Planet"
output:
  rmarkdown::html_document:
    code_folding: hide
    theme: sandstone   
    highlight: tango
    css: "css/andrewl.css"
    includes:
      before_body: "partials/header.html"
      after_body: "partials/footer.html"
---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> Want to see the code? Click on the black boxes on the right to show/hide the code. 


### Map

Water makes up nearly three-quarters of the Earth's surface. It's an amazing thing really
- if the atmosphere was any heavier, at a higher pressure the seas would boil away, and if
our planet's orbit was slightly further away from the sun the seas would be frozen.

This animation shows countries disapearing and it's area added to an ever-growing circle. The map
is an equal-area map which gives you a good idea of just how blue our planet is.

```{r data, message=FALSE, warning=FALSE, results='show'}
#First, we'll need to load a bunch of libraries so we can handle and view geospatial data
library(sf)
library(ggplot2)
library(rnaturalearth)
library(gganimate)
library(purrr)
library(dplyr)

sf::sf_use_s2(TRUE)

#Load the shapefile for the world boundaries
a <- st_read('./data/11/TM_WORLD_BORDERS-0.2.shp', quiet = TRUE)

#Transform the shapefile to Mollweide projection - this is an equal-area projection
b <- a %>% st_transform(crs = "ESRI:54009")

#Add an id column to the data
for(i in 1:nrow(b)) {
  b$id[i] = i
}

# Create sf polygon using WGS84 extent coordinates then project to Mollweide
# This will be the circle that grows as the countries disappear
bbox <- data.frame(lon = c(rep(-180, 180), seq(-180, 180, length.out = 100),
                           rep(180, 180), seq(180, -180, length.out = 100)),
                   lat = c(seq(-90, 90, 1), rep(90, 99), 
                           seq(90, -90, -1), rep(-90, 99))) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON") %>%
  st_transform("ESRI:54009")

# #Loop through each country and create a map
unique(b$id) %>% purrr::walk(function(this_id) {

  #Filter the data to only include countries with an id greater than the current id
  filtered <-  filter(b, b$id>this_id)

  #Calculate the total area of the countries that have been removed
  #and set the area of the circle to be the same
  removed <-  filter(b, b$id<=this_id)
  p <- st_as_sfc("POINT(0 0)")
  st_crs(p) <- 3857
  total_area <- sum(st_area(removed))
  radius <- sqrt(total_area/pi)
  circle <- st_buffer(p,radius)  %>% st_transform(crs = st_crs(b))


  #Plot the map
  ggplot() +
    geom_sf(data = bbox, fill = "darkblue", col = "darkblue", linewidth = 0.25) +
    geom_sf(data = filtered, fill= "lightgreen", color="darkblue") +
    geom_sf(data = circle, fill= "lightgreen", color="darkblue") +
    theme_void() +
    theme(axis.title = element_blank(),
          plot.background = element_rect(fill = "transparent", colour = NA),
          panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(fill = NA,
                                      colour = "#238B45",
                                      linewidth = 2,
                                      inherit.blank = TRUE))

  ggsave(sprintf("/tmp/blue-planet-%i.png", this_id))

})

system('ffmpeg -framerate 30 -i /tmp/blue-planet-%d.png -vf "crop=2100:1044:0:525,scale=900:430" -c:v libx264 -preset slow -crf 23 -pix_fmt yuv420p images/blue-planet.mp4')

```


![](images/blue-planet.mp4)


### Credits

World Boundaries: Natural Earth (Public Domain) https://www.naturalearthdata.com/about/
